### 2014-11-10 Arvind.tex
Pass through speculation

The branch predictor is the simpler of the two components, whose state
is indicated with metavariable $\ppc$.
The operations on such state are $\curPpc(\ppc)$, to extract the
current program-counter prediction; $\nextPpc(\ppc)$, to get the predicted
program-counter for the next instruction; and $\setBp(\ppc, \pc)$,
to reset prediction to begin at a known-accurate position $\pc$. We need to impose no explicit correctness criterion on
branch predictors; the processor uses predictions only as hints, and
it always resets the predictor using $\setBp$ after detecting an
inaccurate hint.

The interface and formal contract of a reorder buffer are more
involved.  We write $\rob$ as a metavariable for reorder buffer
state. $\phi$ denotes the state of an empty buffer. The operations associated with $\rob$ are:
\begin{itemize}
%\item $\phi$, the state of an empty buffer.
\item $\add(\pc, \rob)$, which appends the program instruction at location $\pc$ to the list of instructions that the buffer is allowed to consider executing.
\item $\compute(\rob)$, which models a
step of computation inside the buffer, returning both an updated state
and an optional speculative load to issue. For instance, it invokes the $\dec$
and $\exec$ functions (as defined for SC) internally to obtain the next program
counter, state, \etc{} (but the actual states are not updated).
\item $\upd\ld(\rob,t,v)$, which informs the buffer that the memory
has returned result value $v$ for the speculative load with tag $t \neq \epsilon$.
\item $\commit(\rob)$, which returns the next instruction in serial
program order, if we have accumulated enough memory responses to execute it
accurately, or returns
$\epsilon$ otherwise.  When $\commit$ returns an instruction, it also
returns the associated program counter plus the next program counter
we would advance to afterward.
 Furthermore, the instruction is
extended with any relevant response from memory (used only for load
instructions, obtained through $\upd\ld$) and with the new architectural state (\eg{} register
file) after execution.
\item $\retire(\rob)$, which informs the buffer that its $\commit$
instruction was executed successfully, so it is time to move on to the
next instruction.
\end{itemize}

@@

\inference[LdRpBad]
{\commit(\rob) = (\pc,\pc',(\ld,x,a,v',s')) \\ v' \neq v & \exec(s,\pc,(\ld,x,v)) = (s'',\pc'')}
{\iox{\text{P$_\text{so}$}}{s,\pc,\top,\\\rob,\ppc}{s'',\pc'',\bot,\phi,\\\setBp(\ppc,\pc)}{\text{ToP}(\epsilon, \ld, v)}}

@@

A full processor state is $(s, \pc, \wait, \rob, \ppc)$, comprising
architectural state, the program counter, a Boolean flag indicating
whether the processor is waiting for a memory response about an
instruction being committed, and the reorder buffer and branch
predictor states. Its initial state is given by $(s_0, \pc_0, \bot, \phi, \ppc_0)$.
The interface of this processor with memory (\ie{}
communication labels with $\toM, \toP$) is identical
to that of the reference processor.
% but now we can exercise a memory
%system's transitions for speculative loads.

@@

\begin{corollary}
\label{ges}
$\text{P$^n_\text{so}$} \sqsubseteq_{\noSpec^n}$ P$_\text{ref}^n$
\end{corollary}
\begin{prf}
Direct consequence of Theorems~\ref{ocorrect},~\ref{liftn} (the
latter is one of our preliminary results about $n$-repetition).
\end{prf}

>>>>

The branch predictor is the simpler of the two components, whose state
is indicated with metavariable $\ppc$.
The operations on such state are $\curPpc(\ppc)$, to extract the
current program-counter prediction; $\nextPpc(\ppc)$, to advance to predicting
the next instruction; and $\setBp(\ppc, \pc)$,
to reset prediction to begin at a known-accurate position $\pc$. We need to impose no explicit correctness criterion on
branch predictors; the processor uses predictions only as hints, and
it always resets the predictor using $\setBp$ after detecting an
inaccurate hint.

The interface and formal contract of a reorder buffer are more
involved.  We write $\rob$ as a metavariable for reorder-buffer
state. $\phi$ denotes the state of an empty buffer. The operations associated with $\rob$ are:
\begin{itemize}
%\item $\phi$, the state of an empty buffer.
\item $\add(\pc, \rob)$, which appends the program instruction at location $\pc$ to the list of instructions that the buffer is allowed to consider executing.
\item $\compute(\rob)$, which models a
step of computation inside the buffer, returning both an updated state
and an optional speculative load to issue. For instance, it invokes the $\dec$
and $\exec$ functions (as defined for SC) internally to obtain the next program
counter, state, \etc{} (but the actual states are not updated).
\item $\upd\ld(\rob,t,v)$, which informs the buffer that the memory
has returned result value $v$ for the speculative load with tag $t \neq \epsilon$.
\item $\commit(\rob)$, which returns the next instruction in serial
program order, if we have accumulated enough memory responses to execute it
accurately, or returns
$\epsilon$ otherwise.  When $\commit$ returns an instruction, it also
returns the associated program counter plus the next program counter
we would advance to afterward.
 Furthermore, the instruction is
extended with any relevant response from memory (used only for load
instructions, obtained through $\upd\ld$) and with the new architectural state (\eg{} register
file) after execution.
\item $\retire(\rob)$, which informs the buffer that its $\commit$
instruction was executed successfully, so it is time to move on to the
next instruction.
\end{itemize}

@@

\inference[LdRpBad]
{\commit(\rob) = (\pc,\pc',(\ld,x,a,v',s')) \\ v' \neq v & \exec(s,\pc,(\ld,x,v)) = (s'',\pc'')}
{\iox{\text{P$_\text{so}$}}{s,\pc,\top,\\\rob,\ppc}{s'',\pc'',\bot,\phi,\\\setBp(\ppc,\pc'')}{\text{ToP}(\epsilon, \ld, v)}}

@@

A full processor state is $(s, \pc, \wait, \rob, \ppc)$, comprising
architectural state, the program counter, a Boolean flag indicating
whether the processor is waiting for a memory response about an
instruction being committed, and the reorder-buffer and branch-predictor
states. Its initial state is given by $(s_0, \pc_0, \bot, \phi, \ppc_0)$.
The interface of this processor with memory (\ie{}
communication labels with $\toM, \toP$) is identical
to that of the reference processor.
% but now we can exercise a memory
%system's transitions for speculative loads.

@@

\begin{corollary}
\label{ges}
$\text{P$^n_\text{so}$} \sqsubseteq_{\noSpec^n}$ P$_\text{ref}^n$
\end{corollary}
\begin{prf}
Direct consequence of Theorems~\ref{ocorrect} and \ref{liftn} (the
latter is one of our preliminary results about $n$-repetition).
\end{prf}
